European or American? A re-examination of the transatlantic influence over the IMF

### 2025-07-31 manuscript/manuscript.qmd
correct references

European influence reflects a strategic combination of formal power and coordinated action. EU countries collectively control over 30% of IMF voting rights, and if they were to unify, they could potentially counterbalance US veto power [@brandner2009a; @leech2005a; @smaghi2004]. This is reinforced by the long-standing transatlantic norm allocating IMF leadership to a European national [@kahler2001]. Through platforms such as EURIMF and SCIMF, EU members have harmonised policy positions and influenced loan design. During the Eurozone crisis, France and Germany pushed for IMF involvement to stabilise the region while embedding austerity-oriented conditionality that protected their banking systems [@hodson2014; @kickert2019a; @veron2016]. Lending terms were shaped to reflect creditor priorities, with evidence that conditionality was both intensified or relaxed based on geopolitical alignment [@lutz2014; @heimberger2025, p. 9; @dreher2015]. Beyond Europe, the UK and France have used IMF concessional lending in Africa to maintain post-colonial ties and commercial access [@stone2004].

>>>>

European influence reflects a strategic combination of formal power and coordinated action. EU countries collectively control over 30% of IMF voting rights, and if they were to unify, they could potentially counterbalance US veto power [@brandner2009a; @leech2005a; @smaghi2004]. This is reinforced by the long-standing transatlantic norm allocating IMF leadership to a European national [@kahler2001]. Through platforms such as EURIMF and SCIMF, EU members have harmonised policy positions and influenced loan design. During the Eurozone crisis, France and Germany pushed for IMF involvement to stabilise the region while embedding austerity-oriented conditionality that protected their banking systems [@hodson2014; @kickert2019a; @veron2016]. Lending terms were shaped to reflect creditor priorities, with evidence that conditionality was both intensified or relaxed based on geopolitical alignment [@cain2022; @dreher2015; @lutz2014]. Beyond Europe, the UK and France have used IMF concessional lending in Africa to maintain post-colonial ties and commercial access [@moyo2024; @stone2004].
